---
output: github_document
---

```{r, include = FALSE}
lgr::get_logger("bbotk")$set_threshold("warn")
set.seed(1)
options(datatable.print.class = FALSE, datatable.print.keys = FALSE)
```

# bbotk - Black-Box Optimization Toolkit

Package website: [release](https://bbotk.mlr-org.com/) | [dev](https://bbotk.mlr-org.com/dev/)

<!-- badges: start -->
[![tic](https://github.com/mlr-org/bbotk/workflows/tic/badge.svg?branch=main)](https://github.com/mlr-org/bbotk/actions)
[![CRAN Status Badge](https://www.r-pkg.org/badges/version-ago/bbotk)](https://cran.r-project.org/package=bbotk)
<!-- badges: end -->

This package provides a common framework for optimization including

* `Optimizer`: Objects of this class allow you to optimize an object of the class `OptimInstance`.
* `OptimInstance`: Defines the optimization problem, consisting of an `Objective`, the `search_space` and a `Terminator`.
   All evaluations on the `OptimInstance` will be automatically stored in its own `Archive`.
* `Objective`: Objects of this class contain the objective function.
   The class ensures that the objective function is called in the right way and defines, whether the function should be minimized or maximized.
* `Terminator`: Objects of this class control the termination of the optimization independent of the optimizer.

Various optimization methods are already implemented e.g. grid search, random search and generalized simulated annealing.

## Resources

* Package [vignette](https://CRAN.R-project.org/package=bbotk/vignettes/bbotk.html)

## Installation

Install the last release from CRAN:

```{r eval = FALSE}
install.packages("bbotk")
```

Install the development version from GitHub:

```{r eval = FALSE}
remotes::install_github("mlr-org/bbotk")
```

## Examples

```{r, include = FALSE}
# mute load messages
library(bbotk)
```

### Quick optimization with `bb_optimize`

```{r}
library(bbotk)

# define objective function
fun = function(xs) {
  c(y = - (xs[[1]] - 2)^2 - (xs[[2]] + 3)^2 + 10)
}

# optimize function with random search
result = bb_optimize(fun, method = "random_search", lower = c(-10, -5), upper = c(10, 5), max_evals = 100)

# optimized parameters
result$par

# optimal outcome
result$value
```

### Advanced optimization

```{r}
# define objective function
fun = function(xs) {
  c(y = - (xs[[1]] - 2)^2 - (xs[[2]] + 3)^2 + 10)
}

# set domain
domain = ps(
  x1 = p_dbl(-10, 10),
  x2 = p_dbl(-5, 5)
)

# set codomain
codomain = ps(
  y = p_dbl(tags = "maximize")
)

# create Objective object
objective = ObjectiveRFun$new(
  fun = fun,
  domain = domain,
  codomain = codomain,
  properties = "deterministic"
)

# Define termination criterion
terminator = trm("evals", n_evals = 10)

# create optimization instance
instance = OptimInstanceSingleCrit$new(
  objective = objective,
  terminator = terminator
)

# load optimizer
optimizer = opt("gensa")

# trigger optimization
optimizer$optimize(instance)

# best performing configuration
instance$result

# all evaluated configuration
as.data.table(instance$archive)
```
