---
output: github_document
---

```{r, include = FALSE}
library(bbotk)
lgr::get_logger("bbotk")$set_threshold("warn")
set.seed(1)
options(
  datatable.print.nrows = 10,
  datatable.print.class = FALSE,
  datatable.print.keys = FALSE,
  width = 100)
```

# bbotk - Black-Box Optimization Toolkit

Package website: [release](https://bbotk.mlr-org.com/) | [dev](https://bbotk.mlr-org.com/dev/)

<!-- badges: start -->
[![r-cmd-check](https://github.com/mlr-org/bbotk/actions/workflows/r-cmd-check.yml/badge.svg)](https://github.com/mlr-org/bbotk/actions/workflows/r-cmd-check.yml)
[![CRAN Status Badge](https://www.r-pkg.org/badges/version-ago/bbotk)](https://cran.r-project.org/package=bbotk)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
<!-- badges: end -->

*bbotk* is a black-box optimization framework for R.
It features highly configurable search spaces via the [paradox](https://github.com/mlr-org/paradox) package and optimizes every user-defined objective function.
The package includes several optimization algorithms e.g. Random Search, Grid Search, Iterated Racing, Bayesian Optimization (in [mlr3mbo](https://github.com/mlr-org/mlr3mbo)) and Hyperband (in [mlr3hyperband](https://github.com/mlr-org/mlr3hyperband)).
bbotk is the base package of [mlr3tuning](https://github.com/mlr-org/mlr3tuning), [mlr3fselect](https://github.com/mlr-org/mlr3fselect) and [miesmuschel](https://github.com/mlr-org/miesmuschel).

## Resources

There are several sections about black-box optimization in the [mlr3book](https://mlr3book.mlr-org.com).
Often the sections about tuning are also relevant for general black-box optimization.

* Getting started with [black-box optimization](https://mlr3book.mlr-org.com/chapters/chapter5/advanced_tuning_methods_and_black_box_optimization.html#sec-black-box-optimization).
* An overview of all optimizers and tuners can be found on our [website](https://mlr-org.com/tuners.html).
* Learn about log transformations in the [search space](https://mlr3book.mlr-org.com/chapters/chapter4/hyperparameter_optimization.html#sec-logarithmic-transformations).
* Or more advanced [search space transformations](https://mlr3book.mlr-org.com/chapters/chapter4/hyperparameter_optimization.html#sec-tune-trafo).
* Run [multi-objective optimization](https://mlr3book.mlr-org.com/chapters/chapter5/advanced_tuning_methods_and_black_box_optimization.html#sec-multi-metrics-tuning).
* The [mlr3viz](https://github.com/mlr-org/mlr3viz) package can be used to [visualize](https://mlr-org.com/gallery/technical/2022-12-22-mlr3viz/#tuning-instance) the optimization process.
* Quick optimization with the [`bb_optimize`](https://bbotk.mlr-org.com/reference/bb_optimize.html) function.

## Installation

Install the latest release from CRAN.

```{r eval = FALSE}
install.packages("bbotk")
```

Install the development version from GitHub.

```{r eval = FALSE}
pak::pkg_install("mlr-org/bbotk")
```

## Example

```{r}
# define the objective function
fun = function(xs) {
  list(y = - (xs[[1]] - 2)^2 - (xs[[2]] + 3)^2 + 10)
}

# set domain
domain = ps(
  x1 = p_dbl(-10, 10),
  x2 = p_dbl(-5, 5)
)

# set codomain
codomain = ps(
  y = p_dbl(tags = "maximize")
)

# create objective
objective = ObjectiveRFun$new(
  fun = fun,
  domain = domain,
  codomain = codomain,
  properties = "deterministic"
)

# initialize instance
instance = oi(
  objective = objective,
  terminator = trm("evals", n_evals = 20)
)

# load optimizer
optimizer = opt("gensa")

# trigger optimization
optimizer$optimize(instance)

# best performing configuration
instance$result

# all evaluated configuration
as.data.table(instance$archive)
```

